---
title: 'Reproducible Research Peer-graded Assignment: Course Project 1'
author: "Kevin McCue"
date: "July 17, 2017"
output:
  html_document: 
    keep_md: true
---


## Loading and preprocessing the data
##
##
##1.1 Downloading and loading the full dataset
```{r Part1.1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~")
rm(list = ls())
# load all packages used in this exploratory analysis
library(knitr)
library(dplyr)
library(ggplot2)
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
data <- read.csv(unz(temp, "activity.csv"))
unlink(temp)

```

##1.2 Setting up all the data for analysis
```{r Part1.2, echo = TRUE}
data$dayofweek <- weekdays(as.Date(data$date))
averagestepsperdayofwk <- aggregate(steps ~ weekdays(as.Date(date)), data = data, FUN = mean)
colnames(averagestepsperdayofwk)[which(names(averagestepsperdayofwk) == "weekdays(as.Date(date))")] <- "dayofweek"
sdstepsperdayofwk <- aggregate(steps ~ weekdays(as.Date(date)), data = data, FUN = sd)
colnames(sdstepsperdayofwk)[which(names(sdstepsperdayofwk) == "weekdays(as.Date(date))")] <- "dayofweek"
averagestepsperint <- aggregate(steps ~ interval, data = data, FUN = mean)
sdstepsperdayofint <- aggregate(steps ~ interval, data = data, FUN = sd)
steps.date <- aggregate(steps ~ date, data = data, FUN = sum)
```

##2.  Bar Plot of Number of Steps Per Day
``` {r Part2, fig.width = 10, fig.align = "center", fig.asp = .5, echo=TRUE}
barplot(steps.date$steps, names.arg = steps.date$date, 
        xlab = "Date", 
        ylab = "Step Count", 
        main = paste("Steps per Day From ", min(as.Date(data$date)), " To ", max(as.Date(data$date))))
```
##
##



## What is mean total number of steps taken per day?

## 3.1 The mean number of steps per day:
```{r Part3.1, echo=TRUE}
 mean(steps.date$steps)
```

## 3.2 The median number of steps per day:
```{r Part3.2, echo=TRUE}

median(steps.date$steps)
```

## What is the average daily activity pattern?


##4.  Time series plot of the average number of steps taken

```{r Part4, fig.width = 10, fig.align = "center", fig.asp = .5, echo=TRUE}
plot(averagestepsperint, type = "l", 
     xlab= "Interval", 
     ylab = "Steps", 
     main = paste("Averaged Steps Per Interval From ", min(as.Date(data$date)), " To ", max(as.Date(data$date))))
```

##5.  The 5-minute interval that, on average, contains the maximum number of steps
```{r Part5, echo=TRUE}
averagestepsperint$interval[which.max(averagestepsperint$steps)]
```



## Imputing missing values

##6.  Code to describe and show a strategy for imputing missing data
```{r Part6, echo=TRUE}
sum(is.na(data))
data <- merge(data, averagestepsperint, by = "interval", suffixes = c("", ".y"))
nas <- is.na(data$steps)
data$steps[nas] <- data$steps.y[nas]
data <- data[, c(1:3)]
sum(is.na(data))
```

##7.  Histogram of the total number of steps taken each day after missing values are imputed
```{r Part7, fig.width = 10, fig.align = "center", fig.asp = .5, echo=TRUE}
hist(steps.date$steps, 
        xlab = "Date", breaks = 20, 
     col = "grey",
        ylab = "Step Count", 
        main = paste("Total Number of Steps Taken Each Day Steps per Day From ", min(as.Date(data$date)), " To ", max(as.Date(data$date))))
```



## Are there differences in activity patterns between weekdays and weekends?

##8.  Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r Part8, fig.width = 10, fig.align = "center", fig.asp = .6, echo=TRUE}
daytype <- function(date) {
  if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
    "weekend"} else {"weekday"}
}
data$daytype <- as.factor(sapply(data$date, daytype))
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
  steps.type <- aggregate(steps ~ interval, data = data, subset = data$daytype == 
                            type, FUN = mean)
  plot(steps.type, type = "l", main = type, xlab= "Interval", 
       ylab = "Steps" )
}
```